---
title: "Project 1 (Beer) Markdown"
author: "Michael Schulte"
date: "10/9/2022"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Link to recording: https://smu.zoom.us/rec/share/Gujqwi30bZtZpix1Wlj58pEX7dx37r-WDR0gRYhTein89fP_C1e9dtbBbIxzKQOd.slY31p9lFCdaozCi?startTime=1666477030000
Passcode: FST@4!kH

Load in data
```{r}
library(tidyverse)
beers <- data.frame(read.csv('Beers.csv', stringsAsFactors = FALSE))
breweries <- data.frame(read.csv('Breweries.csv', stringsAsFactors = FALSE))
```
1.	How many breweries are present in each state?
```{r}
num_breweries <- breweries %>% ggplot(aes(x = State, fill = State)) + geom_bar()+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
geom_text(stat='count', aes(label=..count..), vjust=-1)
num_breweries
```
Colorado has the largest amount of breweries.

2.	Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.  (RMD only, this does not need to be included in the presentation or the deck.)
```{r}
head(beers)
head(breweries)
breweries <- breweries %>% rename("Brewery_id" = "Brew_ID")
bb <- merge(beers, breweries, by="Brewery_id", all=TRUE)
```
```{r}
bb <- bb %>% rename("Beer" = "Name.x") %>% rename("Brewery" = "Name.y")
```
```{r}
print(head(bb))
print(tail(bb))
```
3.	Address the missing values in each column.
```{r}
summary(bb)
# bb[is.na(bb)] <- 0
library(plyr)
impute.mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
bb <- ddply(bb, ~ Style, transform, ABV = impute.mean(ABV), IBU = impute.mean(IBU))
bb[is.na(bb)] <- 0
```
```{r}
summary(bb)
```
4.	Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.
```{r}
library(ggplot2)
bb %>% ggplot(aes(x = as.factor(State), y = ABV, fill = State)) + geom_bar(stat = "summary", fun = "median") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ggtitle("Median ABV")

bb %>% ggplot(aes(x = as.factor(State), y = IBU, fill = State)) + geom_bar(stat = "summary", fun = "median") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ggtitle("Median IBU")
```
ME has both the highest median ABV and the highest median IBU of all the states. There is a lot of overlap between median ABV and median IBU.

5.	Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
```{r}
paste0("Max ABV State:", bb$State[which.max(bb$ABV)])
bb %>% slice_max(ABV)
paste0("Max IBU State:", bb$State[which.max(bb$IBU)])
```
The max ABV states is Colorado, and the max IBU state is Oregon.

6.	Comment on the summary statistics and distribution of the ABV variable.
```{r}
summary(bb$ABV)
bb %>% ggplot(aes(x = ABV)) + geom_histogram() + ggtitle("Histogram of ABV")
```
The min is .1%, the max is 12.8%, and the median is 5.65%. This shows that most Americans prefer beer than is around the 5%-6% ABV.
7.	Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.  Make your best judgment of a relationship and EXPLAIN your answer.
```{r}
bb %>% ggplot(aes(x = IBU, y = ABV)) + geom_smooth(method = "lm") + geom_point() + ggtitle("ABV vs IBU")
```
The linear regression line shows a clear correlation between ABV and IBU. As ABV increases, so does IBU.

8.	Budweiser would also like to investigate the difference with respect to IBU and ABV between IPAs (India Pale Ales) and other types of Ale (any beer with “Ale” in its name other than IPA).  You decide to use KNN classification to investigate this relationship.  Provide statistical evidence one way or the other. You can of course assume your audience is comfortable with percentages … KNN is very easy to understand conceptually. 
In addition, while you have decided to use KNN to investigate this relationship (KNN is required) you may also feel free to supplement your response to this question with any other methods or techniques you have learned.  Creativity and alternative solutions are always encouraged.
```{r}
unique(bb$Style)
bb$Class1 <- NA
for(i in 1:nrow(bb)){
  if(grepl("IPA", bb$Style[i]) == TRUE | grepl("India Pale Ale", bb$Style[i]) == TRUE){
    bb$Class1[i] <- "IPA"
  } else if(grepl("Ale", bb$Style[i]) == TRUE & !grepl("India Pale Ale", bb$Style[i]) == TRUE & !grepl("IPA", bb$Style[i]) == TRUE){
    bb$Class1[i] <- "Ale"
  } else{
    bb$Class1[i] <- NA
  }
}
```
```{r}
library(class)
library(caret)
library(e1071)
set.seed(1234)
bb_knn_df <- na.omit(bb)
splitPerc <- 0.80
trainIndices <- sample(1:dim(bb_knn_df)[1], round(splitPerc * dim(bb_knn_df)[1]))
train <- bb_knn_df[trainIndices,]
test <- bb_knn_df[-trainIndices,]
train$ABV <- scale(train$ABV)
train$IBU <- scale(train$IBU)
test$ABV <- scale(test$ABV)
test$IBU <- scale(test$IBU)
accuracy <- c()
k <- c()
# Find maximum accuracy and its respective k value
for(i in 1:300){
  classifications <- knn(train[, c(4, 5)], test[, c(4, 5)], train$Class1, prob = TRUE, k = i)
  CM <- confusionMatrix(table(classifications, test$Class1))
  accuracy[i] <- CM$overall[1]
  k[i] <- i
  if(i == 1){
    max_k <- i
    accuracy_max <- accuracy[i]
  }else if(accuracy[i] >= accuracy_max){
    max_k <- i
    accuracy_max <- accuracy[i]
  }
}
plot(k, accuracy, type = "l", xlab = "k")
print(paste0("Max k:", max_k))
print(paste0("Max accuracy:", accuracy[max_k]))
```
```{r}
classifications <- knn(train[, c(4, 5)], test[, c(4, 5)], train$Class1, prob = TRUE, k = max_k)
  CM <- confusionMatrix(table(classifications, test$Class1))
  print(CM)
```
The accuracy is 0.8893. This is rather good for a KNN model. The sensitivity and specificity are also high. Also, the pos pred value is high meaning that there are a lot of true values that were predicted.
```{r}
library(plotly)
x_test <- test %>% select("ABV", "IBU")
y_test <- test %>% select("Class1")
yscore <- knn(train[, c(4, 5)], test[, c(4, 5)], train$Class1, prob = TRUE, k = max_k)
yscore <- attributes(yscore)$prob

pdb <- cbind(x_test, y_test)
pdb <- cbind(pdb, yscore)

fig <- plot_ly(data = pdb,x = ~IBU, y = ~ABV, type = 'scatter', mode = 'markers',color = ~yscore, colors = 'RdBu', symbol = ~Class1, split = ~Class1, symbols = c('square-dot','circle-dot'), marker = list(size = 12, line = list(color = 'black', width = 1)))
fig
```
The above plot shows the divisions between the IPAs and Ales graphically. The index shows the probability/the intesisty which the beer was categorized.
```{r}
p3<-ggplot(train,aes(x=IBU,y=ABV,colour=Class1)) + geom_jitter() +geom_density2d() + ggtitle("Density Plot of ABV vs IBU")
p3
```
  

9.	Knock their socks off!  Find one other useful inference from the data that you feel Budweiser may be able to find value in.  You must convince them why it is important and back up your conviction with appropriate statistical evidence. 
```{r}
fit.IBU <- aov(IBU~Style, data = bb)
print("ANOVA IBU")
summary(fit.IBU)
fit.ABV <- aov(ABV~Style, data = bb)
print("ANOVA ABV")
summary(fit.ABV)
```
There is significant evidence to prove that at least one Style has a different IBU, and one style has a different ABV (p-value < 2e-16).As a result, one can infer that different styles are characterized by different combinations of bitterness and alcohol level. The plot below shows that there are different ABV and IBU levels for each style of beer. More so, there appears to be a greater correlation between style and IBU than style and ABV.
```{r}
bb %>% ggplot(aes(x = IBU, y = ABV, color = Style)) + geom_point() + theme(legend.position = "None") + ggtitle("ABV vs IBU by Style")
```